Figure supplements

### 2022-08-21 text/Manuscript.Rmd
Upload Supplement Figures

Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology. In nervous systems, NO can modulate circuits, but its effects on neural dynamics and behaviour remain unclear. Here we show how NO signalling shapes circuit dynamics to orchestrate light-avoidance behaviour. We studied UV-avoidance mediated by brain ciliary photoreceptor cells (cPRCs) in larval *Platynereis dumerilii*, a marine annelid. We found NOS specifically expressed in interneurons (INNOS) postsynaptic to cPRCs. Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production. NO feeds back to cPRCs and triggers delayed cPRC activation through an unconventional guanylyl cyclase. This results in the activation of a second population of postsynaptic interneurons (INRGW). In *NOS* knockout larvae, NO feedback and INRGW activation do not occur and avoidance behaviour is defective. Our results reveal how NO-mediated neuroendocrine signalling gates a synaptic circuit to initiate light-avoidance behaviour.

@@

We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data. Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences <!---# need to add a NOS tree as supplement -->. To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis. In two-day-old larvae, we detected *NOS* expression in two cells in the apical organ region (Fig. 1A) <!---# 3d HCR and transgene? -->. The same two cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene. This reporter also revealed the axonal projections of the NOS-expressing neurons. In three-day-old larvae... <!---# need 3d HCR and transgene data as well -->

@@

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively. (C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (A) and (B) are shown in purple. Other neurons are shown in light grey. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D and E) Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at 2-day-old. (D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). (E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (F) Wiring diagram of the cPRC circuit. Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively. Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins.](../figures/Fig2.png){alt="Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance. (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins." width="100%"}

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light. Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells. (#Kei - are there time course data that could be shown?). We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low. These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 2 - figure supplement ). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs. In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants. This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.**  (A-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1. (B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (C) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. The white dashed lines in (C) and (D) indicate the outer frame of the larvae. (E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (G) Co-localisation analysis using NIT-GC1 and NOS antibodies. (B-G) Anterior view of the 2-day-old larvae.NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig3.png){width="100%"}

@@

@@

![**Figure 4. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead). A black arrowhead indicates the position of the adult eye. The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs. (C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm). Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean. (D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae. (E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae. (F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR. White arrowheads indicate INNOS strongly expressing RYa-pNP. (G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line). (H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae. (I) Schematic diagram of the immunostaining procedure after calcium imaging. (J) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (K) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (J). INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown. (L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig4.png){width="100%"}

>>>>

Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology. In nervous systems, NO can modulate circuits, but its effects on neural dynamics and behaviour remain unclear. Here we show how NO signalling shapes circuit dynamics to orchestrate light-avoidance behaviour. We studied UV-avoidance mediated by brain ciliary photoreceptor cells (cPRCs) in larval *Platynereis dumerilii*, a marine annelid. We found NOS specifically expressed in interneurons (INNOS) postsynaptic to cPRCs. Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production. NO feeds back to cPRCs and triggers delayed cPRC activation through an unconventional guanylyl cyclase. This results in the activation of a second population of postsynaptic interneurons (INRGW). It was also suggested that activation of INRGWs activates serotonergic ciliomotor neurons (Ser-h1), which are postsynaptic and involved in increasing ciliary beat frequency. In *NOS* knockout larvae, NO feedback and INRGW activation do not occur and avoidance behaviour is defective. Our results reveal how NO-mediated neuroendocrine signalling gates a synaptic circuit to initiate light-avoidance behaviour.

@@

We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data. Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Fig. 1---figure supplement 1). To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis. In two-day-old larvae, we detected *NOS* expression in two cells in the apical organ region (Fig. 1A) <!---# 3d HCR, Kei will put them -->. The same two cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene. This reporter also revealed the axonal projections of the NOS-expressing neurons. In three-day-old larvae... <!---# need 3d HCR , Kei will put them -->

@@

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 72 hpf Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR. Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). Anterior view of the larva at 54 and 72 hpf. (E) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 72 hpf from the anterior side. (F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 72 hpf larva. The INNOS on the ventral side labelled with (E) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres.  (H) Wiring diagram of the cPRC circuit. <!---# This graph widget does not look very good. We need to edit this a bit more. --> Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively. Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins.](../figures/Fig2.png){alt="Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance. (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins." width="100%"}

@@

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (E and F) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae. (G-J) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n > 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig3.png){width="100%"}

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light. Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells. We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low. These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 1). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs. In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants. This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.

@@

![**Figure 4. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead). A black arrowhead indicates the position of the adult eye. The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs. (C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm). Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean. (D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae. (E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae. (F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR. White arrowheads indicate INNOS strongly expressing RYa-pNP. (G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line). (H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae. (L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig4.png){width="100%"}

### 2022-08-21 text/Supplements.Rmd
Upload Supplement Figures

![**Figure 1—figure supplement 1. (A) Phylogenetic tree of NOS by maximum likelihood (ML). Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}
![**Figure 3—figure supplement 1. (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}

>>>>

---
title: Figure supplements
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*'
output:
  pdf_document:
    latex_engine: lualatex
    template: default
  html_document:
    theme: lumen
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

![**Figure 1---figure supplement 1.** (A) Phylogenetic tree of NOS by maximum likelihood (ML). Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}

![**Figure 1---figure supplement 2.** (A) Immunostaining with NOS antibodies.(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva.](../figures/Fig1_sup2.png){width="100%"}

![\*\*Figure 3---figure supplement 1.\*\* (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}

![**Figure 3---figure supplement 2.** (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. (B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig3_sup2.png){width="100%"}

![**Figure 3---figure supplement 3.** (A) Comparison of amino acid sequences of regions of possible NO binding in NIT-GC1 and NIT-GC2. (B) 3D stuructures of NIT-GC1 and NIT-GC2 (C)](../figures/Fig3_sup3.png){width="100%"}

![**Figure 3---figure supplement 4.** (A and B) Left: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Right shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence.](../figures/Fig3_sup4.png){width="100%"}

![**Figure 4---figure supplement 1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging. (B) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (C) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (B). INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown. .](../figures/Fig4_sup1.png){width="100%"}
